Management induced changes of soil organic carbon on global croplands

### 2020-10-06 SOCbudgetPaper/SOCbudget.tex
Update SOCbudget.tex

\begin{abstract}
Soil organic carbon (SOC) is one of larges carbon stocks on earth. It is
three times larger than the biospheric pool and more than twice as big
as the athmospheric pool, when looking into the first meter of the earth
soil profile. Human cropping activties led and still lead to a depletion
of SOC though, which are so far not well represented in global
assessments of historic carbon emissions. While SOC models often
represent well the biochemical processes that lead to the accumulation
and decay of SOC, the management decisions driving these biophysical
processes are still little investigated on the global scale. Here we
create a spatial explicit data set for agricutural management on
cropland, especially for crop residue and manure management, based on
global historic production (FAOSTAT), and land-use (LUH2) data; and
combine it with the IPCC Tier 2 approach to create a half degree
resolution SOC stock changes on mineral soils. We estimate that due to
arable farming soils have lost over 37 GtOC of which 4 GtOC have been
regained within the period 1975-2010. We show that, our results on
global scale based on Tier 2 IPCC methodolgy are in good agreement with
Tier 1 default assumptions. We also find that SOC is very sensitive to
management decision such as residue recycling indicating the nessessity
to incorporated better management data in soil models.
\end{abstract}

@@

\begin{figure}[H]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas. (b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades. (c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics. Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas.}\label{fig:SOCmaps}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario} \caption{Global SOC gap in GtC for various stylized management counterfactual scenarios compared to the modeled historical baseline.}\label{fig:SOCscen}
\end{figure}

\begin{figure}[H]
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/OuFlowFig} \caption{Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues. Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil).}\label{fig:FlowFig}
\end{figure}

@@

This study provides an analysis on historic SOC stock changes on
cropland. We determine the SOC trends on cropland compared to a
potential natural vegeted world under the same historic climatic
conditions (\(SOC_{natveg}\)) leading to a cummulative SOC emission
since dawn of mankind of around 37 GtC in 2010. Whereas recent modelling
estimates on global SOC emissions indicate an on going increase of SOC
emissions (Sanderman et al, Pugh et al.), our study indicates that the
global SOC gap might be closing slowly.

According to Sanderman et al.~cummulative SOC emissions since beginning
of human cropping activities have been at around 37 GtC for the first 30
cm of the soil with half of it attributed to grazing. It was also
pointed out, that this results might be conservative estimates and low
compared to experimental results, leaving our estimate of 37 GtC in 2010
for cropland emissions only in alignment with Sanderman et al.~estimates
considering the hugh uncertainties in modelling SOC on the global scale.
Furtherore the results of Sanderman et al.~calculated historical trends
based on agricultural land expansion without considering SOC variations
due to different management systems at all. In Pugh et al.~management
effects like tillage and residue recycling have been considered, but
neither changes over time nor alignment to oberserved historical data
like yields levels or no tillage areas were taken into account. Under
this assumptions this study only found marginal effect of crop
productivity and other management effects on cummulative SOC emissions.

Nevertheless our results show, when purely focusing on the past decades,
the moderate global cropland expansion of around 11\% between 1974 and
2010, is outweighted by improved agricultural yields ad practises.
Moreover our sensitivity analysis indicades that 1. yield increases and
with that the increase in residue biomass might play the most important
role, followed by 2. enhanced residue recycling rates, 3. improved
manure recycling (e.g.~due to improved animal waste management system)
and 4. the adoption of no tillage practises.

Modelling management effects on the global scale comes however with huge
parametic and systematic uncertainties. One aspect pointed out by Keel
(2017) as well as by Smith (2019) might be that carbon input
calculations are highly sensitive to the choice of allometric functions
determining below and above ground residue estimates from harvested
quantities. Keel et al.~questioned, that below ground residues might
increase with a fixed root:shot ratio rather than being independend from
productivity gains. Following this argumentation SOC results shown in
this study might especially in high-yielding farming system overestimate
actual SOC stocks. However according to our study above ground residue
biomass recycling seem to contribute even more to overall SOC stocks due
to higher input rates.

\subsection{Modeled management effect inline with default IPCC
assumptions}

@@

Our climate zone specific, aggregated \(F_{SCF}\) correspond very well
to the tier 1 estimation of IPCC, 2006 of \(F_{SOC}\). For the tropical
regions the assumptions changed notablly from the guidelines in 2006 to
the update in 2019, leaving our results too low in comparison with IPCC,
2019. Considering yield gaps in mainly developing regions in the tropics
the default assumption of medium input systems, might be an
overestimation of actual SOC state. The additional effect of considering
low inpout regimes in tropical regions can however not explain the full
mismatch to IPCC 2019 values but account for at least 5-7\% of it.

>>>>

\begin{abstract}
Soil organic carbon (SOC) is one of larges carbon stocks on Earth. It is
three times larger than the biospheric pool and more than twice as big
as the athmospheric pool, when looking into the first meter of the earth
soil profile. Human cropping activties led and still lead to a depletion
of SOC though, which are so far not well represented in global
assessments of historic carbon emissions. While SOC models often
represent well the biochemical processes that lead to the accumulation
and decay of SOC, the management decisions driving these biophysical
processes are still little investigated on the global scale. Here we
create a spatial explicit data set for agricutural management on
cropland, especially for crop residue and manure management, based on
global historic production (FAOSTAT), and land-use (LUH2) data; and
combine it with the IPCC Tier 2 approach to create a half degree
resolution SOC stock changes on mineral soils. We estimate that due to
arable farming soils have lost over 37 GtOC of which 4 GtOC have been
regained within the period 1975-2010. We show that, our results on
global scale based on Tier 2 IPCC methodolgy are in good agreement with
Tier 1 default assumptions. We also find that SOC is very sensitive to
management decision such as residue recycling indicating the nessessity
to incorporated better management data in soil models.
\end{abstract}

@@

\begin{figure}[H]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks on cropland shows high carbon stocks in high yielding areas. (b): The SOC debt is decreasing over time, meaning net SOC gains on global croplands over the last decades. (c)+(d): Absolute (c) and relative (d) SOC stocks compared to a potential natural state showing different hot spots of SOC dynamics. Whereas the absolute losses might be in temperate dry regions, relative losses are more prominent in tropical moist areas. (BB: Increase map size, legend is larger than map, no need for that)}\label{fig:SOCmaps}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/scenario} \caption{Global SOC gap in GtC for various stylized management counterfactual scenarios compared to the modeled historical baseline. (BB: order of legend should be equal to order of lines. Management=tillage+reside+manure?)}\label{fig:SOCscen}
\end{figure}

\begin{figure}[H]
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/OuFlowFig} \caption{Global carbon flows (small numbers) and stocks (bold numbers) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues. Note the two numbers on carbon inputs to soil denote carbon applied to the field and carbon entering the soil (difference is decomposted before official counting as soil).(BB: adjuste flow size to numbers, include arrows for losses, e.g. --> x ,otherwise its confusing, trees need black outer line, roots have no clean upper end)}\label{fig:FlowFig}
\end{figure}

@@

This study provides an analysis of historic SOC stock changes on
cropland. We determine the SOC trends on cropland compared to a counterfactual scenario with a world under natural vegetation under identical historical climatic
conditions (\(SOC_{natveg}\)). Our results show, that human activity lead to cummulative SOC emissions
of around 37 GtC in 2010. Whereas recent modelling
estimates of global SOC emissions indicate an ongoing increase of SOC
emissions (Sanderman et al, Pugh et al.), our study indicates that the
global SOC gap is slowly closing due to improved management.

According to Sanderman et al.~cummulative SOC emissions since beginning
of human cropping activities have been at around 37 GtC for the first 30
cm of the soil with half of it attributed to grazing. It was also
pointed out that these results might be conservative estimates and low
compared to experimental results. Considering the huge uncertainties in modelling SOC on the global scale, 
our estimate of 37 GtC in 2010 for cropland emissions only still seems consistent with Sanderman et al.~estimates.

Furthermore (BB: consider using a spelling checker) the results of Sanderman et al.~calculated (BB: the results calculated?) historical trends
based on agricultural land expansion without considering SOC variations
due to different management systems at all. Pugh et al.~ considered management
effects like tillage and residue recycling in a static way, but
neither changes over time nor alignment to oberserved historical data
like yields-levels or no-tillage areas were taken into account. Under
these assumptions this study only found marginal effects (BB: be more precise) of crop
productivity and other management effects on cummulative SOC emissions.

Our study for the first time uses a dynanmic management dataset as driver for SOC dynamics. We
show that the moderate global cropland expansion of around 11\% between 1974 and
2010 and the resulting depletion of SOC stocks in converted cropland has been outweighted by improved agricultural yields ad practises.
Moreover our sensitivity analysis indicades that (1) yield increases and
with that the increase in residue biomass might play the most important
role, followed by (2) enhanced residue recycling rates, (3) improved
manure recycling (e.g.~due to improved animal waste management system)
and (4) the adoption of no tillage practises.

Modelling management effects on the global scale comes however with 
parametric and structural uncertainties. As pointed out by Keel
(2017) and Smith (2019), carbon input
calculations are highly sensitive to the choice of allometric functions
determining below and above ground residue estimates from harvested
quantities. Keel et al.~questioned whether below ground residues might
increase with a fixed root:shoot ratio rather than being independend from
productivity gains. Following this argumentation SOC results shown in
this study might especially in high-yielding farming system overestimate
actual SOC stocks (BB: No. Very likely, root:shoot ratios come from developed countries. They rather underestimate SOC in LICs). However (?), according to our study above ground residue biomass recycling seems to contribute even more to overall SOC stocks due
to higher input rates.

\subsection{Modeled management effect in line with default IPCC
assumptions}

@@

Our climate zone specific, aggregated \(F_{SCF}\) correspond very well
to the tier 1 estimation of IPCC, 2006 of \(F_{SOC}\). For the tropical
regions the assumptions changed notably from the guidelines in 2006 to
the update in 2019, leaving our results too low in comparison with IPCC,
2019. Considering yield gaps in mainly developing regions in the tropics
the default assumption of medium input systems, might be an
overestimation of actual SOC state. The additional effect of considering
low inpout regimes in tropical regions can however not explain the full
mismatch to IPCC 2019 values but account for at least 5-7\% of it.
